---
title: "Data Visualization Lab"
author: "Wai Lam Adele Hong A15999023"
date: "10/13/2021"
output: pdf_document
---

Install the package ggplot2.
```{r}
# install.packages("ggplot2")
```
Anytime I want to use this package, I need to load it.
```{r}
library(ggplot2)
```

Quick base R Plot - Cars
```{r}
View(cars)
plot(cars)
# This is NOT ggplot. It's only a simple plot.
```

Let's make our first ggplot. We need data + aes + geoms (data, aesthetic, geometry).
```{r}
ggplot(data=cars) + aes(x=speed, y=dist) + geom_point()
# "aes" is x and y axes. & visuals. "geom" is type of plot (how it's plotted, e.g point, line, bar, etc.)

p <- ggplot(data=cars) + aes(x=speed, y=dist) + geom_point()

# Add a line with geom_line
p + geom_line()
p + geom_smooth()
# Code below shows trendline ("lm" means linear model)
p + geom_smooth(method="lm")
```

--------------------------------------------
Read in our drug expression data
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

Q. How many genes are there?
```{r}
nrow(genes)
```
Q. How many columns and what are the names?
```{r}
ncol(genes)
colnames(genes)
```

Q. How many "up" regulated genes?
```{r}
table(genes$State)
# For a summary of data frame, use table() function
```

Q. Using your values above and 2 significant figures, what fraction of total genes is up regulated in this dataset?
```{r}
round((table(genes$State) / nrow(genes)) * 100, 2)
# Each value for State divided by the total # of genes. The "2" at the end indicates how many decimals  to round it to.
```

Let's make our first plot attempt.
NOTE: if we used the variable "p", it would override previous command regarding "p" (storing obj in "p")
```{r}
g <- ggplot(data=genes) + aes(x=Condition1, y=Condition2, col=State) +geom_point()
# col=State color-codes based on types of State.

g

# Change colors
g + scale_color_manual(values=c("blue", "gray", "red")) +
  labs(title="Gene expression changes", x="Control (no drug)", y="Drug Treatment") + theme_bw()
# With "labs() function, we can add title, axis titles, legends, etc.
# theme_bw function changes theme of graph
```

--------------------------------------------

**Optional Section**

Install gapminder package, which is dataset for economic and demographic info of various countries throughout the yrs.
```{r}
# install.packages("gapminder")
library(gapminder)
gapminder
```

Install **dplyr code** to focus in on a single year.
```{r}
# install.packages("dplyr")
library(dplyr)

gapminder_2007 <- gapminder %>% filter(year==2007)
# Filters to contain only the rows with a year value of 2007
```

Let's make a scatterplot for 2007 data subset.
```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point(alpha=0.5)
# "alpha=0.5" helps make data points a little more transparent to see overlap more clearly
```

Add variables to aes, using additional arguments.
```{r}
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.5)
# Each dimension adds a new layer to the plot! Cool!
```

What happens if we display "pop" using color?
```{r}
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, color = pop) +
  geom_point(alpha=0.8)
# Since the "pop" data is continuous (as opposed to discrete), it is displayed as a color *spectrum*
```

When we use size to display "pop", we realize the sizes don't proportionally reflect the population (i.e the sizes are binned by default). To fix this, we use **scale_size_area()** function.
```{r}
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, size = pop) +
  geom_point(alpha=0.5) + 
  scale_size_area(max_size = 10)
```

Q. Can you adapt your plot for the year 1957? 
A. Yes. It's relatively difficult to compare the 2007 plot to the 1957 because they have different scales (i.e x and y axes have different ranges).
```{r}
# Don't forget to filter yr 1957 and save to "gapminder_1957"
gapminder_1957 <- gapminder %>% filter(year==1957)
ggplot(gapminder_1957) + 
  aes(x = gdpPercap, y = lifeExp,
                 size = pop) +
  geom_point(alpha=0.7) + 
  scale_size_area(max_size = 15)
```
Q. Let's compare them in an easier way; put both years as an input for ggplot. Use **facet_wrap()** function.
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)

ggplot(gapminder_1957) + 
  geom_point(aes(x = gdpPercap, y = lifeExp, color=continent,
                 size = pop), alpha=0.7) + 
  scale_size_area(max_size = 10) +
  facet_wrap(~year)
# Notice geom_point could also be written like this ^
```

```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point()
# Note: this is also a viable way of writing the code above:
ggplot(gapminder, aes(gdpPercap, lifeExp)) +
  geom_point()
```

