---
title: "Lab10: Halloween candy!"
author: "Wai Lam Adele Hong A15999023"
date: "10/28/2021"
output: html_document
---

# Importing Candy Data

Here we look at 538 Halloween candy data. 
```{r}
url <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"
candy <- read.csv(url, row.names=1)
head(candy, n=5)
# "winpercent"= when people are given the choice between 2 candy, how often does the person pick the given candy.
```

**Q1.** How many different candy types are in this dataset?
```{r}
nrow(candy)
# 85 candy types.
```

**Q2.** How many fruity candy types are in the dataset?
```{r}
# "0" means candy is not fruity, and "1" means candy is fruity.
table(candy$fruity)
# or
sum(candy$fruity)
# 38 fruity candy!
```

# What is your favorite candy?
```{r}
candy["Welch", ]$winpercent
# or
candy["Welch", "winpercent"]
```

**Q3.** What is your favorite candy in the dataset and what is it’s winpercent value?

My favorite candy is Welch Fruit Snacks, and the winpercent value is 44.38.

**Q4.** What is the winpercent value for “Kit Kat”?
```{r}
candy["Kit Kat", "winpercent"]
# winpercent = 76.77
```

**Q5.** What is the winpercent value for “Tootsie Roll Snack Bars”?
```{r}
candy["Tootsie Roll Snack Bars", "winpercent"]
# winpercent = 49.65
```

```{r}
# install.packages("skimr")
library("skimr")
skim(candy)
# "skim()" function gives you a quick overview of a given dataset. Let’s install this package and try it on our candy data.
```

**Q6.** Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

The winpercent looks like it's on a different scale.

**Q7.** What do you think a zero and one represent for the candy$chocolate column?

A zero means that the candy is not a chocolate, and a one means that it is a chocolate.

Let's make a histogram.

**Q8.** Plot a histogram of winpercent values.
```{r}
hist(candy$winpercent)
```

**Q9.** Is the distribution of winpercent values symmetrical?

No, the distribution of winpercent values are slightly skewed right.

**Q10.** Is the center of the distribution above or below 50%?

The center of distribution is below 50%.

**Q11.** On average is chocolate candy higher or lower ranked than fruit candy?
```{r}
inds <- as.logical(candy$chocolate)
chocolate <- candy[inds,]$winpercent
inds.fruit <- as.logical(candy$fruity)
fruity <- candy[inds.fruit,]$winpercent

mean(chocolate)
mean(fruity)
```
Chocolate candy is higher ranked than fruit candy.

**Q12.** Is this difference statistically significant?
```{r}
t.test(chocolate, fruity)
```
Yes, the differentce is statiscally significant because the t-test has a small p-value.

# Overall Candy Rankings

**Q13.** What are the five least liked candy types in this set?
```{r}
head(candy[order(candy$winpercent), ], n=5)
# Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble, and Jawbusters are the least liked candies.
```

**Q14.** What are the top 5 all time favorite candy types out of this set?

```{r}
head(candy[order(candy$winpercent, decreasing=TRUE), ], n=5)
# ReeseÕs Peanut Butter cup, ReeseÕs Miniatures, Twix, Kit Kat, and Snickers are top5 favorite candies.
```

**Q15.** Make a first barplot of candy ranking based on winpercent values.
```{r}
library("ggplot2")

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col() +
  theme(text = element_text(size = 6))
```

**Q16.** This is quite ugly, use the reorder() function to get the bars sorted by winpercent?
```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col() +
  theme(text = element_text(size = 6))
```
Let's add some useful color.
We start by making a vector of all black values (one for each candy). Then we overwrite chocolate (for chocolate candy), brown (for candy bars) and red (for fruity candy) values.
```{r}
# Make a color vector
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

# Apply color vector
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill = my_cols) +
  theme(text = element_text(size = 6))
```



**Q17.** What is the worst ranked chocolate candy?

The worst ranked chocolate candy is Charleston Chew.



# Taking a look at pricepercent

What about value for money? What is the the best candy for the least money? Solution: Make a plot of winpercent vs the pricepercent variable.
```{r}
ggplot(candy) +
  aes(winpercent, pricepercent) +
  geom_point()
```

Add labels so we know which point is which candy.
```{r}
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point() +
  geom_text()
```

Let's add some useful color.
We start by making a vector of all black values (one for each candy). Then we overwrite chocolate (for chocolate candy), brown (for candy bars) and red (for fruity candy) values.
```{r}
# Make a color vector
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
# Apply color vector
ggplot(candy) +
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_point(fill=my_cols)
```





